---
title: "VHA SAIL Metrics Datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{VHA Sail metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)

devtools::load_all()

```


VA SAIL metrics are a way of summarizing VA hospital performance data.  Individual data sets by location are available for download from [VA.gov](https://www.va.gov/QUALITYOFCARE/measure-up/Strategic_Analytics_for_Improvement_and_Learning_SAIL.asp).  These are excel datasets formatted with banners and special characters.

When provided a file's url, `sail_tidy` downloads, imports, and wrangles the dataset into a format that's easier to utilize in R.

```{r}
library(tidyVA)
library(ggplot2)
library(dplyr)


albany_sail <- sail_tidy("https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_albany_fy19q1.xlsx")

glimpse(albany_sail)

```

Using `map_df` from the __purrr__ package you can create a data frame of metrics from several sites:

```{r}

albany <-  "https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_albany_fy19q1.xlsx"
syracuse <- "https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_syracuse_fy19q1.xlsx"
bath <- "https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_bath_fy19q1.xlsx"
buffalo <- "https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_buffalo_fy19q1.xlsx"
bronx <- "https://www.va.gov/QUALITYOFCARE/SAIL_FY19_Q1/SAIL-2_bronx_fy19q1.xlsx"

links <- c(albany, syracuse, bath,
           buffalo,bronx)

data <- purrr::map_df(links, sail_tidy)

```

This can then easily be explored within the __tidyverse__

```{r}
data %>% 
  distinct(Measure)

```

```{r}

data %>%
  filter(Measure =="1. Acute care Standardized Mortality Ratio (SMR)") %>%
  mutate(SMR = as.numeric(Value)) %>% 
  ggplot(aes(reorder(Site, SMR), SMR)) +
  geom_point() +
  geom_segment(aes(x = Site, y = 0, 
                   xend = Site, yend = SMR)) +
  coord_flip() + 
  geom_hline(yintercept = 0.409, lty = "dashed", color = "darkred") +
  labs(x = NULL, y = "Standardized Ratio") +
  ggtitle("Acute Care Standardized Mortality Ratio (SMR)",
          subtitle = "VISN 2 - 2019 Q1") +
  theme_va(grid = FALSE)
  
```




```{r}

data %>% 
  filter(Measure == "1. Population coverage") %>%
  mutate(Value = as.numeric(Value)) %>%
  arrange(desc(Value)) %>% 
  ggplot(aes(forcats::fct_inorder(Site), Value)) +
  geom_col(alpha = .75, fill = "#fdbf11", width = .75 ) +
  labs(y = "Standardized Score", x = "Year-Quarter",
       color= "Site") +
  ggtitle("Population Coverage - Mental Health",
          subtitle = "SAIL Metrics for VISN 2") +
  theme_va(grid = "XY")
```



```{r}

data %>% 
  filter(Measure == "1. Population coverage") %>%
  mutate(value = as.numeric(Value),
         best_10 = as.numeric(`Best 10%`)) %>% 
  arrange(desc(value)) %>% 
  ggplot(aes(forcats::fct_inorder(Site), value)) +
  geom_point() +
  geom_segment(aes(x = Site, y = best_10, 
                   xend = Site, yend = value)) +
  coord_flip()  +
  geom_hline(aes(yintercept = best_10), 
             lty = "solid", color = "#fdbf11") +
  labs(x = NULL, y = "Standardized Score") +
  ggtitle("VISN 2 Mental Health Coverage",
          subtitle = "Distance from Best 10%") +
  theme_va(grid = "Y") 

``` 




